Web-based Supplementary Materials for A Space-time \Skewt Model for Threshold Exceedances by Morris, Reich, Thibaud, and Cooley

### 2016-06-28 LaTeX/supplemental_rev.tex
updates to manuscript

\title[Web-based Supplementary Materials for A Space-time Skew-$t$ Model for Threshold Exceedances]{Web-based Supplementary Materials for A Space-time Skew-$t$ Model for Threshold Exceedances by Morris, Reich, Thibaud, and Cooley}
\author
{Samuel A Morris$^{1,*}$\email{samorris@ncsu.edu},
Brian J Reich$^{1}$,
Emeric Thibaud$^{2}$, and
Daniel Cooley$^{2}$\\
$^{1}$Department of Statistics, North Carolina State University, Raleigh, North Carolina, U.S.A. \\
$^{2}$Department of Statistics, Colorado State University, Fort Collins, Colorado, U.S.A.}

\begin{document} %\linenumbers
\maketitle
% \begin{center}
% {\Large {\bf Supplemental material for A space-time skew-$t$ model for threshold exceedances}}\\

@@

\section{Skew-$t$ distribution} \label{sta:skewt}
\subsection*{Univariate skew-$t$ distribution}
We say that $Y$ follows a univariate extended skew-$t$ distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, and degrees of freedom $\nu$ if has distribution function
\begin{align}
  f_{\text{EST}}(y) = 2 f_T (z; \nu) F_T\left[ \alpha z \sqrt{ \frac{ \nu + 1 }{ \nu + z^2}}; \nu + 1 \right]
\end{align}
where $f_T(t; \nu)$ is a univariate Student's $t$ with $\nu$ degrees of freedom, $F_T(t; \nu) = P(T < t)$, and \hbox{$z = (y - \xi) / \omega$}.

\subsection*{Multivariate skew-$t$ distribution}
If $\bZ \sim \text{ST}_d(0, \bar{\bOmega}, \balpha, \eta)$ is a $d$-dimensional skew-$t$ distribution, and $\bY = \xi + \bomega \bZ$, where $\bomega = \text{diag}(\omega_1, \ldots, \omega_d)$, then the density of $Y$ at $y$ is
\begin{align}
  f_y(\by) = det(\bomega)^{-1} f_z(\bz)
\end{align}
where
\begin{align}
  f_z(\bz) &= 2 t_d(\bz; \bar{\bOmega}, \eta) T \left[ \balpha^\top \bz \sqrt{ \frac{\eta + d}{\nu + Q(\bz)} }; \eta + d\right] \\
  \bz &= \bomega^{-1}(\by - \xi)
\end{align}
where $t_d(\bz; \bar{\bOmega}, \eta)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\eta$, $Q(z) = \bz^\top \bar{\Omega}^{-1}\bz$ and $T(\cdot; \eta)$ denotes the univariate Student's $t$ distribution function with $\eta$ degrees of freedom \citep{Azzalini2014}.

\subsection*{Extremal dependence}
For a bivariate skew-$t$ random variable $\bY = [Y(\bs), Y(\bt)]^\top$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
\begin{align} \label{steq:chiskew-t}
  \chi(h) = \bar{F}_{\text{EST}}\left\{ \frac{[x_1^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{\sqrt{1 - \varrho(h)^2}}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \frac{ [x_2^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{ \sqrt{1 - \varrho(h)^2} }; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\},
\end{align}
where $\bar{F}_{\text{EST}}$ is the univariate survival extended skew-$t$ function with zero location and unit scale, \hbox{$\varrho(h) = \text{cor}[y(\bs), y(\bt)]$}, $\alpha_j = \alpha_i \sqrt{1 - \varrho^2}$, $\tau_j = \sqrt{\eta + 1}(\alpha_j + \alpha_i \varrho)$, and $x_j = F_T(\bar{\alpha}_i \sqrt{\eta + 1}; 0, 1, \eta) / F_T(\bar{\alpha}_j \sqrt{\eta + 1}; 0, 1, \eta)$ with $j = 1, 2$ and $i = 2, 1$ and where $\bar{\alpha}_j = (\alpha_j + \alpha_i \varrho) / \sqrt{ 1 + \alpha_i^2 [1 - \varrho(h)^2]}$.

\subsection*{Proof that $\lim_{h \rightarrow \infty} \chi(h) > 0$}
Consider the bivariate distribution of $\bY = [Y(\bs), Y(\bt)]^\top$, with $\varrho(h)$ given by \eref{steq:matern}.
So, $\lim_{h \rightarrow \infty} \varrho(h) = 0$.
Then
\begin{align}
  \lim_{h \rightarrow \infty} \chi(h) = \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\}.
\end{align}
Because the extended skew-$t$ distribution is not bounded above, for all $\bar{F}_{\text{EST}}(x) = 1 - F_{\text{EST} (x)} > 0$ for all $x < \infty$.
Therefore, for a skew-$t$ distribution, $\lim_{h \rightarrow \infty} \chi(h) > 0$.

\section{Comparisons with other parameterizations} \label{sta:otherparams}
Various forms of multivariate skew-normal and skew-$t$ distributions have been proposed in the literature.
In this section, we make a connection between our parameterization in \eref{steq:fullmodel} of the main text and another popular version.
\citet{Azzalini2014} and \citet{Beranger2016} define a skew-normal process as
\begin{align}
  \tilde{X}(\bs) = \tilde{\lambda}|z| + (1 - \tilde{\lambda}^2)^{1 / 2} v(\bs)
\end{align}
where $\tilde{\lambda} \in (-1, 1)$, $z \sim N(0, 1)$, and $v(\bs)$ is a Gaussian process with mean zero, variance one, and spatial correlation function $\rho$.
To extend this to the skew-$t$ distribution, \citet{Azzalini2003} take $\tilde{Y}(\bs) = W\tilde{X}(\bs)$ where $W^{-2} \sim $ Gamma$(a / 2, a / 2)$.
Returning to the proposed parameterization (with $\bbeta = 0$), let $W^{-2} = \frac{b}{a}\sigma^{-2} \sim$ Gamma$(a / 2, a / 2)$ so that \eref{steq:fullmodel} in the manuscript becomes
\begin{align}
  Y(\bs) = W \left[ \lambda \left(\frac{b}{a}\right)^{1 / 2} |z| + \left(\frac{b}{a}\right)^{1 / 2} v(\bs) \right].
\end{align}
Clearly setting $b = a (1 - \tilde{\lambda}^2) > 0$, and $\lambda = \tilde{\lambda} / (1 - \tilde{\lambda}^2)^{1 / 2} \in (-\infty, \infty)$ resolves the difference in parameterizations.
We note that our parameterization has three parameters $(a, b, \lambda)$ compared to the two parameters of the alternative parameterization $(a, \tilde{\lambda})$.
Since we have assumed that both $v(\bs)$ and $z$ have unit scale, the additional $b$ parameter in our parameterization is required to control the precision.
% However, if we introduce an overall scale parameter $c > 0$ into the alternative parameterization so that $\tilde{Y}(\bs) = c W \tilde{X}(\bs)$, then the two models remain equivalent by setting $a = \nu$, $b = \frac{\nu}{c^2 (1 - \tilde{\lambda}^2)}$, and $\lambda = \tilde{\lambda} / (1 - \tilde{\lambda}^2)^{1 / 2}$.

@@

\section{Brier scores for ozone prediction} \label{sta:ozonesite}
Because typical ozone concentration varies throughout the US, we have included Brier scores by site for two model fits (Gaussian and Skew-$t$, $K = 5$, $T = 50$) in \fref{stfig:bssite}.
As we can see in these plots, both models seem to perform similarly across the US with the poorest performance in California.
Other methods have similar Brier score maps to these.
\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{plots/bs-site}
  \caption{Map of Brier scores for Gaussian (top) vs Skew-$t$, $K = 5$, $T = 50$ (bottom).}
  \label{stfig:bssite}
\end{figure}

@@

\begin{table}[htbp]
  \centering
  \caption{Setting 2 -- Skew-$t$ marginal, $K = 1$ knot}
  \label{sttbl:st1sim}
  \begin{tabular}{|l|cccc|cccc|cccc|ccc|}
    \cline{2-16}
    \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
    \hline
    Method 1 &   & B &   &   &   & B &   &   &   & B &   &   &   & B &   \\
    \hline
    Method 2 & A &   &   &   & A &   &   &   & A &   &   &   & A &   &   \\
    \hline
    Method 3 & A & B &   &   & A & B &   &   & A & B &   &   & A & B &   \\
    \hline
    Method 4 & A & B &   &   & A & B &   &   & A & B &   &   & A & B &   \\
    \hline
    Method 5 &   &   & C &   &   &   & C &   &   &   & C &   &   &   & C \\
    \hline
    Method 6 &   &   &   & D &   &   &   & D &   &   &   & D &   &   & C \\
    \hline
  \end{tabular}
\end{table}

\begin{table}[htbp]
  \centering
  \caption{Setting 3 -- Skew-$t$ marginal, $K = 5$ knots}
  \label{sttbl:st5sim}
  \begin{tabular}{|l|cccc|cccc|ccc|ccc|}
    \cline{2-15}
    \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
    \hline
    Method 1 &   &   & C &   &   &   & C &   &   & B &   &   & B &   \\
    \hline
    Method 2 &   &   & C &   &   &   & C &   &   & B &   &   & B &   \\
    \hline
    Method 3 &   & B &   &   &   & B &   &   & A &   &   & A &   &   \\
    \hline
    Method 4 & A &   &   &   & A &   &   &   & A &   &   & A &   &   \\
    \hline
    Method 5 & A &   &   &   & A &   &   &   & A &   &   & A &   &   \\
    \hline
    Method 6 &   &   &   & D &   &   &   & D &   &   & C &   &   & C \\
    \hline
  \end{tabular}
\end{table}

>>>>

\title[Web-based Supplementary Materials for A Space-time \Skewt Model for Threshold Exceedances]{Web-based Supplementary Materials for A Space-time \Skewt Model for Threshold Exceedances by Morris, Reich, Thibaud, and Cooley}
\author
{Samuel A Morris$^{1,*}$\email{samorris@ncsu.edu},
Brian J Reich$^{1}$,
Emeric Thibaud$^{2}$, and
Daniel Cooley$^{2}$\\
$^{1}$Department of Statistics, North Carolina State University, Raleigh, North Carolina, U.S.A. \\
$^{2}$Department of Statistics, Colorado State University, Fort Collins, Colorado, U.S.A.}

\begin{document} %\linenumbers
\maketitle
% \begin{center}
% {\Large {\bf Supplemental material for A space-time \skewt model for threshold exceedances}}\\

@@

\section{\Skewt distribution} \label{sta:skewt}
\subsection*{Univariate \skewt distribution}
We say that $Y$ follows a univariate extended \skewt distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, and degrees of freedom $\nu$ if has distribution function
\begin{align}
  f_{\text{EST}}(y) = 2 f_T (z; \nu) F_T\left[ \alpha z \sqrt{ \frac{ \nu + 1 }{ \nu + z^2}}; \nu + 1 \right]
\end{align}
where $f_T(t; \nu)$ is a univariate Student's $t$ with $\nu$ degrees of freedom, $F_T(t; \nu) = P(T < t)$, and \hbox{$z = (y - \xi) / \omega$}.

\subsection*{Multivariate \skewt distribution}
If $\bZ \sim \text{ST}_d(0, \bar{\bOmega}, \balpha, \eta)$ is a $d$-dimensional \skewt distribution, and $\bY = \xi + \bomega \bZ$, where $\bomega = \text{diag}(\omega_1, \ldots, \omega_d)$, then the density of $Y$ at $y$ is
\begin{align}
  f_y(\by) = det(\bomega)^{-1} f_z(\bz)
\end{align}
where
\begin{align}
  f_z(\bz) &= 2 t_d(\bz; \bar{\bOmega}, \eta) T \left[ \balpha^\top \bz \sqrt{ \frac{\eta + d}{\nu + Q(\bz)} }; \eta + d\right] \\
  \bz &= \bomega^{-1}(\by - \xi)
\end{align}
where $t_d(\bz; \bar{\bOmega}, \eta)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\eta$, $Q(z) = \bz^\top \bar{\Omega}^{-1}\bz$ and $T(\cdot; \eta)$ denotes the univariate Student's $t$ distribution function with $\eta$ degrees of freedom \citep{Azzalini2014}.

\subsection*{Extremal dependence}
For a bivariate \skewt random variable $\bY = [Y(\bs), Y(\bt)]^\top$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
\begin{align} \label{steq:chiskew-t}
  \chi(h) = \bar{F}_{\text{EST}}\left\{ \frac{[x_1^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{\sqrt{1 - \varrho(h)^2}}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \frac{ [x_2^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{ \sqrt{1 - \varrho(h)^2} }; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\},
\end{align}
where $\bar{F}_{\text{EST}}$ is the univariate survival extended \skewt function with zero location and unit scale, \hbox{$\varrho(h) = \text{cor}[y(\bs), y(\bt)]$}, $\alpha_j = \alpha_i \sqrt{1 - \varrho^2}$, $\tau_j = \sqrt{\eta + 1}(\alpha_j + \alpha_i \varrho)$, and $x_j = F_T(\bar{\alpha}_i \sqrt{\eta + 1}; 0, 1, \eta) / F_T(\bar{\alpha}_j \sqrt{\eta + 1}; 0, 1, \eta)$ with $j = 1, 2$ and $i = 2, 1$ and where $\bar{\alpha}_j = (\alpha_j + \alpha_i \varrho) / \sqrt{ 1 + \alpha_i^2 [1 - \varrho(h)^2]}$.

\subsection*{Proof that $\lim_{h \rightarrow \infty} \chi(h) > 0$}
Consider the bivariate distribution of $\bY = [Y(\bs), Y(\bt)]^\top$, with $\varrho(h)$ given by \eref{steq:matern}.
So, $\lim_{h \rightarrow \infty} \varrho(h) = 0$.
Then
\begin{align}
  \lim_{h \rightarrow \infty} \chi(h) = \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\}.
\end{align}
Because the extended \skewt distribution is not bounded above, for all $\bar{F}_{\text{EST}}(x) = 1 - F_{\text{EST} (x)} > 0$ for all $x < \infty$.
Therefore, for a \skewt distribution, $\lim_{h \rightarrow \infty} \chi(h) > 0$.

\section{Comparisons with other parameterizations} \label{sta:otherparams}
Various forms of multivariate skew-normal and \skewt distributions have been proposed in the literature.
In this section, we make a connection between our parameterization in \eref{steq:fullmodel} of the main text and another popular version.
\citet{Azzalini2014} and \citet{Beranger2016} define a skew-normal process as
\begin{align}
  \tilde{X}(\bs) = \tilde{\lambda}|z| + (1 - \tilde{\lambda}^2)^{1 / 2} v(\bs)
\end{align}
where $\tilde{\lambda} \in (-1, 1)$, $z \sim N(0, 1)$, and $v(\bs)$ is a Gaussian process with mean zero, variance one, and spatial correlation function $\rho$.
To extend this to the \skewt distribution, \citet{Azzalini2003} take $\tilde{Y}(\bs) = W\tilde{X}(\bs)$ where $W^{-2} \sim $ Gamma$(a / 2, a / 2)$.
Returning to the proposed parameterization (with $\bbeta = 0$), let $W^{-2} = \frac{b}{a}\sigma^{-2} \sim$ Gamma$(a / 2, a / 2)$ so that \eref{steq:fullmodel} in the manuscript becomes
\begin{align}
  Y(\bs) = W \left[ \lambda \left(\frac{b}{a}\right)^{1 / 2} |z| + \left(\frac{b}{a}\right)^{1 / 2} v(\bs) \right].
\end{align}
Clearly setting $b = a (1 - \tilde{\lambda}^2) > 0$, and $\lambda = \tilde{\lambda} / (1 - \tilde{\lambda}^2)^{1 / 2} \in (-\infty, \infty)$ resolves the difference in parameterizations.
We note that our parameterization has three parameters $(a, b, \lambda)$ compared to the two parameters of the alternative parameterization $(a, \tilde{\lambda})$.
Since we have assumed that both $v(\bs)$ and $z$ have unit scale, the additional $b$ parameter in our parameterization is required to control the precision.
% However, if we introduce an overall scale parameter $c > 0$ into the alternative parameterization so that $\tilde{Y}(\bs) = c W \tilde{X}(\bs)$, then the two models remain equivalent by setting $a = \nu$, $b = \frac{\nu}{c^2 (1 - \tilde{\lambda}^2)}$, and $\lambda = \tilde{\lambda} / (1 - \tilde{\lambda}^2)^{1 / 2}$.

@@

\section{Brier scores for ozone prediction} \label{sta:ozonesite}
Because typical ozone concentration varies throughout the US, we have included Brier scores by site for two model fits (Gaussian and \Skewt, $K = 5$, $T = 50$) in \fref{stfig:bssite}.
As we can see in these plots, both models seem to perform similarly across the US with the poorest performance in California.
Other methods have similar Brier score maps to these.
\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{plots/bs-site}
  \caption{Map of Brier scores for Gaussian (top) vs \Skewt, $K = 5$, $T = 50$ (bottom).}
  \label{stfig:bssite}
\end{figure}

@@

\begin{table}[htbp]
  \centering
  \caption{Setting 2 -- \Skewt marginal, $K = 1$ knot}
  \label{sttbl:st1sim}
  \begin{tabular}{|l|cccc|cccc|cccc|ccc|}
    \cline{2-16}
    \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
    \hline
    Method 1 &   & B &   &   &   & B &   &   &   & B &   &   &   & B &   \\
    \hline
    Method 2 & A &   &   &   & A &   &   &   & A &   &   &   & A &   &   \\
    \hline
    Method 3 & A & B &   &   & A & B &   &   & A & B &   &   & A & B &   \\
    \hline
    Method 4 & A & B &   &   & A & B &   &   & A & B &   &   & A & B &   \\
    \hline
    Method 5 &   &   & C &   &   &   & C &   &   &   & C &   &   &   & C \\
    \hline
    Method 6 &   &   &   & D &   &   &   & D &   &   &   & D &   &   & C \\
    \hline
  \end{tabular}
\end{table}

\begin{table}[htbp]
  \centering
  \caption{Setting 3 -- \Skewt marginal, $K = 5$ knots}
  \label{sttbl:st5sim}
  \begin{tabular}{|l|cccc|cccc|ccc|ccc|}
    \cline{2-15}
    \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
    \hline
    Method 1 &   &   & C &   &   &   & C &   &   & B &   &   & B &   \\
    \hline
    Method 2 &   &   & C &   &   &   & C &   &   & B &   &   & B &   \\
    \hline
    Method 3 &   & B &   &   &   & B &   &   & A &   &   & A &   &   \\
    \hline
    Method 4 & A &   &   &   & A &   &   &   & A &   &   & A &   &   \\
    \hline
    Method 5 & A &   &   &   & A &   &   &   & A &   &   & A &   &   \\
    \hline
    Method 6 &   &   &   & D &   &   &   & D &   &   & C &   &   & C \\
    \hline
  \end{tabular}
\end{table}
